\documentclass[12pt,a4paper]{report}
\usepackage{graphicx}
\usepackage[color=purple!80]{todonotes} %% does the to do stuff
\usepackage{enumitem}
\usepackage{bbm}	%% makes \1 look good
\usepackage{hyperref}
\usepackage{mathrsfs}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}





\theoremstyle{plain}
\newtheorem{thm}{Theorem}
\newtheorem*{thm*}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem*{lemma*}{Lemma}
\newtheorem*{claim}{Claim}
\newtheorem{prop}{Proposition}
\newtheorem*{prop*}{Proposition}
\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem*{defn*}{Definition}

\newcommand{\1}{\mathbbm{1}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\BH}{\mathcal{\B(\H)}}
\newcommand{\Hr}{\mathcal{H}_\rho}
\newcommand{\Vr}{V_\rho}
\newcommand{\Lr}{L_\rho}
\newcommand{\xr}{x_\rho}
\renewcommand{\phi}{\varphi}
\newcommand{\CX}{C(X)}
\renewcommand{\S}{\mathscr{S}}
\renewcommand{\P}[1]{\mathscr{P}(#1)}
\newcommand{\spec}[1]{\sigma(#1)}
\renewcommand{\labelenumi}{(\roman{enumi})} %% makes top layer enum numbers roman


\author{Luke Armitage}
\title{$C^\ast$-algebras, and the~Gelfand-Naimark~Theorems}

\begin{document}
\maketitle
\makeatletter  						%%todonotes magic
    \providecommand\@dotsep{5}
  \makeatother
\listoftodos\relax
\newcommand\Item[1][]{ 				%% enumerate magic
  \ifx\relax#1\relax  \item \else \item[#1] \fi
  \abovedisplayskip=0pt\abovedisplayshortskip=0pt~\vspace*{-\baselineskip}}
\tableofcontents

\chapter{Preliminaries}

\section{History of The Study of Operator Algebras}
	The noncommutative nature of Werner Heisenberg's work in 1925 on a new quantum mechanics \cite{heisenberg25} 
	led to Born and Jordan \cite{bornjordan25}, together with Heisenberg \cite{bornjordanheisenberg25}, 
	developing the matrix mechanics required to concisely summarise the new quantum mechanical model. 
	From 1935-1943, John von Neumann, together with F.J. Murray, developed the theory of \emph{rings of 
	operators} acting on a Hilbert space \cite{vonneumann35,vonneumann37,vonneumann43,vonneumann40}, in an 
	attempt to establish a general framework for this matrix mechanics.
	These rings of operators are now considered part of the theory of \emph{von Neumann algebras}, a 
	subsection of $C^\ast$-algebra theory. 
	Discussion of the seminal quantum mechanical works of Heisenberg can be found in \cite{mackinnon77}, and 
	similarly \cite{schroer03} gives a summary of the works of Jordan expanding on this.
	
	In 1943 \cite{gelfand43}, Gelfand and Naimark established an abstract characterisation of 
	$C^\ast$-algebras, free from dependence on the operators acting on a Hilbert space.
	The Gelfand-Naimark theorem, which we will be considering here at length, gives the link between these 
	abstract $C^\ast$-algebras and the rings of operators previously studied.
	Used in the proof of the GN theorem is the Gelfand-Naimark-Segal construction, a pair of results relating 
	cyclic $\ast$-representations of $C^\ast$-algebras to certain linear functionals on that algebra. 
	
\section{Background Mathematics and Resources.}	
	The following is some mathematics which may prove useful throughout the project, with relevant resources; 
	we will of course be making definitions as needed, this is for further background and related theory.

	We will be assuming some familiarity with the following theory, giving some explanation as necessary:
\begin{itemize}
	\item Rings, algebras and linear spaces.
	\item Normed spaces, inner product spaces, Banach and Hilbert spaces.
	\item Point-set topology.
\end{itemize}
	A good broad background on all of these can be found in \cite{simmons83}.

	Some texts which cover $C^\ast$-algebras: 
	Dixmier \cite{dixmier77} presents a summary of the general theory up to that time (1977), with 
	\cite{dixmier81} focussing on reworking and developing the theory of von Neumann algebras. 
	Sakai \cite{sakai71} gives a treatment of $C^\ast$- and von Neumann algebras from a more topological point 
	of view. In \cite{kadison83,kadison86}, the authors aim to make accessible the ``vast recent research 
	literature'' in this subsection of functional analysis.
	Blackadar \cite{blackadar06} gives a much faster, more encyclopaedic coverage of the theory of operator 
	algebras, and covering more specialised material and applications.
	
\section{Aims}
	The aims for this project are:
\begin{itemize}
	\item Give a good background understanding on $C^\ast$-algebras, including topological and geometric interpretation of results where possible.
	\item Consider the representation theory of $C^\ast$-algebras, using the Gelfand-Naimark-Segal construction as a starting point.
	\item Consider the commutative and general versions of the Gelfand-Naimark theorem, and understand their contents and proof.
\end{itemize}




most texts start from $\BH$ to justify the whole thing. we're algebraists, who don't need no justification. we jump right in at the deep (abstract) end.
\todo[inline]{rewrite this!} %%
\todo[inline]{summarise entire project chapter by chapter.} %%


\chapter{Foundational definitions and results}


The following notation applies throughout:
$\mathbb{K}$ denotes a field. $R^+$ denotes the non-negative reals. An ideal of an algebra is, unless otherwise stated, taken to mean a two-sided ideal. 

A \emph{partial order} on a set $X$ is a relation $\leq$ which satisfies the conditions
\begin{itemize}
	\item $x \leq x$ for all $x$ in $X$; 		\hfill (reflexivity)
	\item for all $x$ and $y$ in $X$, if $x \leq y$ and $y\leq x$, then $x=y$; 
												\hfill (antisymmetry)
	\item for all $x,y$ and $z$ in $X$, if $x \leq y$ and $y\leq z$, then $x\leq z$. 
												\hfill (transitivity)
\end{itemize}

\section{Normed linear spaces}

Given a linear space $X$ over a field $\mathbb K$, a \emph{linear functional} on $X$ is a linear map 
$\rho:X\to\mathbb K$. The set $X^\ast$ of linear functionals on $X$ is itself a linear space, called
the \emph{dual space} of $X$.

A normed linear space $(V,\|\cdot\|)$ is a \emph{Banach space} if it is complete with respect to 
$\|\cdot\|$, in the sense that all Cauchy sequences in $V$ converge with respect to the norm. An inner 
product space $(V,\langle\cdot,\cdot\rangle)$ is a \emph{Hilbert space} if it is a Banach space with 
respect to the norm induced by $\langle\cdot,\cdot\rangle$.

\begin{thm*}[Hahn-Banach Extension theorem] 
	If $\rho_0$ is a bounded linear functional on a subspace $X_0$ of a normed linear space $X$,
	then there is a bounded linear functional $\rho$ on $X$ such that $\|\rho\|=\|\rho_0\|$ and 
	$\rho=\rho_0$ on $X_0$.
\end{thm*}
\begin{proof}
	Can be found in \cite[Theorem 1.6.1, p.~44]{kadison83}
\end{proof}

\section{Constructions on Hilbert spaces}\label{section:hscons}
We need some Hilbert spaces constructions; particularly, the direct sum of a collection of Hilbert 
spaces and the direct sum of bounded operators on these Hilbert spaces.
Given a finite collection $\{\H_1,\dots,\H_n\}$ of Hilbert spaces, let~$\H$ denote the set
\[
	\H  = \{(x_1,\dots,x_n) ~|~ x_i \in \H_i \mbox { for } i=1,\dots,n\}.
\]
Define addition and scalar multiplication coordinatewise, and given $x=(x_1,\dots,x_n)$ and 
$y=(y_1,\dots,y_n) \in \H$, the equation
\[
	\langle x,y\rangle = \langle x_1,y_1\rangle + \dots + \langle x_n,y_n \rangle
\]
defines an inner product $\langle\cdot,\cdot\rangle$ on $\H$. The resulting norm $\|\cdot\|$ is 
given by
\[
	\|x\|^2 = \|x_1\|^2 + \dots + \|x_n\|^2.
\]
It is easy to show that $\H$ is a Hilbert space with these operations, and we call it the
\emph{direct sum} of the collection $\{\H_1,\dots,\H_n\}$, denoted $\oplus\H_i$.

Similarly, we can construct a Hilbert space direct sum of an infinite collection 
$\{\H_i ~|~ i\in I\}$ of Hilbert spaces. Let $\H$ be the set
\[
	\H= \{(x_i) ~|~ x_i \in \H_i \mbox{ for each }i 
									\mbox{ and } \sum_{i\in I}{\|x_i\|^2} < \infty \}.
\]
Given $x=(x_i)$ and $y=(y_i)\in\H$, we have that
\begin{align*}
			\Big(\sum_{i\in I}{\|x_i+y_i\|^2}\Big)^{1/2}
	&\leq	\Big(\sum_{i\in I}{(\|x_i\|+\|y_i\|)^2\|}\Big)^{1/2}			\\
	&\leq	\Big(\sum_{i\in I}{\|x_i\|^2}\Big)^{1/2}		+
							\Big(\sum_{i\in I}{\|y_i\|^2}\Big)^{1/2}		\\
	&< \infty.
\end{align*}
Hence, the sequence $(x_i+y_i)$ is in $\H$, and we can define addition and scalar multiplication 
coordinatewise on $\H$:
\begin{align*}
	(x_i)+(y_i)=(x_i+y_i) && \alpha(x_i) = (\alpha x_i).
\end{align*}
We also have
\begin{align*}
			\sum_{i\in I}{|\langle x,y\rangle|} 
	&\leq	\sum_{i\in I}{\|x_i\|\|y_i\|}									\\
	&\leq 	\Big(\sum_{i\in I}{\|x_i\|^2}\Big)^{1/2}
							\Big(\sum_{i\in I}{\|y_i\|^2}\Big)^{1/2}		\\
	&< \infty,
\end{align*}
so that we can define an inner product $\langle\cdot,\cdot\rangle$, with induced norm $\|\cdot\|$, 
on $\H$ by
\begin{align*}
	\langle x,y\rangle = \sum_{i\in I}{|\langle x,y\rangle|}, 
					&& \|x\| = \Big(\sum_{i\in I}{\|x_i\|^2}\Big)^{1/2}.
\end{align*}
To see that $\H$ is complete with respect to $\|\cdot\|$, suppose that $(x^{n})_{n\in\N}$ is a 
Cauchy sequence in $\H$, where $x^{n} = (x^n_i)_{i\in I}$ for each $n$. Then given any positive 
$\epsilon$, there exists a positive integer $N$ such that
\[
	\|x^m-x^n\| < \epsilon \mbox{ for all } m,n\geq N,
\]
that is,
\begin{align}\label{eqn:sum1}
	\sum_{i\in I}{\|x^m_i-x^n_i\|^2} < \epsilon^2 \mbox{ for all } m,n\geq N.
\end{align}
Hence for each $i\in I$,
\[
	\|x^m_i-x^n_i\| < \epsilon \mbox{ for all } m,n\geq N
\]
so that $(x^n_i)_{n\in\N}$ is a Cauchy sequence in $\H_i$, having a limit $x_i\in\H_i$.
For any finite subset $J\subset I$, it follows from \eqref{eqn:sum1} that 
\[
	\sum_{j\in J}{\|x^m_j-x^n_j\|^2} < \epsilon^2  \mbox{ for all } m,n\geq N,
\]
and letting $m$ tend to infinity,
\begin{align}\label{eqn:sum2}
	\sum_{j\in J}{\|x_j-x^n_j\|^2} < \epsilon^2  \mbox{ for all } n\geq N.
\end{align}
This holds for any finite subset $J$, so
\begin{align*}
	\sum_{i\in I}{\|x_i-x^n_i\|^2} < \epsilon^2  \mbox{ for all } n\geq N,
\end{align*}
and so $(x_i-x^n_i)$ and $x^n_i$ are in $\H$ for $n\geq N$. 
Then $(x_i)$ is in $\H$ and by \eqref{eqn:sum2}, $x^n$ converges to $(x_i)$ as $n$ tends 
to infinity. We conclude that $\H$ is complete and therefore a Hilbert space. Just like in the finite 
case, we call $\H$ the \emph{direct sum} of the collection $\{\H_i ~|~ i\in I\}$ of Hilbert spaces, 
denoted $\oplus\H_i$.

Suppose now we have a (finite or infinite) collection of bounded operators 
$\{T_i\in\B{(\H_i)} ~|~ i\in I\}$ such that 
\[
	\sup_{i\in I} {\|T_i\|} < \infty
\] 
(by convention this is true when $I$ is a finite set).
For $x=(x_i)$ in $\oplus\H_i$, define an element $Tx$ in $\oplus\H_i$ by $Tx=(T_ix_i)$. 
Then $T:\H\to\H:x\mapsto Tx$ is a bounded linear operator, called the \emph{direct sum} of the 
collection $\{T_i\in\B{(\H_i)} ~|~ i \in I\}$, denoted $\oplus T_i$. For $S_i,T_i$ in 
$\B(\H_i)$, $\alpha,\beta$ in $\C$, we have 
\begin{align*}
		\left(\oplus T_i\right)^\ast &= \oplus T_i^\ast,		\\
		\oplus (\alpha S_i+\beta T_i) &= 
					\alpha \oplus S_i + \beta \oplus T_i, 		\\
		\oplus (S_i T_i) &= \oplus S_i \oplus T_i,				\\
		\|\oplus T_i\| &= \sup_{i\in I} {\|T_i\|}.				\\
\end{align*}

\section{Topological stuff}
\todo[inline]{define: hausdorff/compact/locally compact/denseness}

A \emph{topological vector space} is a vector space $V$ over $\mathbb K$, together with a topology on 
$V$ such that the vector space operations $V\times V\to V:(x,y)\mapsto x+y$ and $\mathbb{K}\times V\to 
V:(\lambda,x)\mapsto \lambda x$ are continuous. 
Recall that an \emph{extreme point} of a convex subset $X_0$ of a topological vector space is a point 
$x$ for which an expression
\begin{align*}
	x = \alpha x_1 +(1-\alpha)x_2,
\end{align*}
for $0\leq\alpha\leq1$ and $x_1,x_2\in X_0$, implies that $x_1=x=x_2$. 
The convex set $X_0$ is then equal to the set of all linear combinations
\begin{align*}
	\alpha_1 x_1+\dots +\alpha_n x_n
\end{align*}
of its extreme points, where $\alpha_1,\dots,\alpha_n$ are positive scalars summing to 1.
For example, for a polygon embedded in $\R^2$, the vertices of a polygon are its extreme points, and
every point within the polygon can be written as a linear combination of the vertices.

\begin{defn*}[{\cite[3.14]{rudin91}}]
	Let $X$ be a topological vector space (over a field $\mathbb K$), with dual space $X^\ast$. Every 
	$x$ in $X$ induces a linear functional $f_x$ on $X^\ast$ defined by $f_x(\rho)=\rho(x)$. The 
	\emph{weak$^\ast$ topology} on $X^\ast$ is the topology generated by the sets 
	\begin{align*}
		\{f^{-1}_x(V) ~|~ x \in X, V\subseteq \mathbb K \mbox{ open}\}.
	\end{align*}
	This is the weakest topology on $X$ such that each functional $f_x$ is continuous, and the 
	collection $\{f_x ~|~ x\in X\}$ separate the points of $X^\ast$, so the weak$^\ast$ topology is 
	Hausdorff.
\end{defn*}
\begin{thm*}[{The Banach-Alaoglu theorem \cite{rudin91}}]
	If $V$ is a neighbourhood of $0$ in a topological vector space $X$, and
	\begin{align*}
		K=\{\rho\in X^\ast ~|~ |\rho(x)|\geq 1 \mbox{ for all } x \in V \},
	\end{align*}
	then $K$ is compact in the weak$^\ast$ topology.
\end{thm*}

\chapter{$C^\ast$-algebras}
We begin this chapter with some definitions and an example, then give some results we 
will need later and finish with a fundamental example.
\begin{defn}
	A \emph{Banach algebra} is a complex Banach space $A$ with norm $\|\cdot\|$ which forms an 
	algebra, such that 
	\begin{align*}
		\|ab\| \leq \|a\| \|b\| \mbox{ for all } a,b \in A.
	\end{align*}
	A \emph{$\ast$-algebra} is an algebra $A$ with an \emph{involution} map 
	$a \mapsto a^\ast$ on $A$ such that, for all $a,b \in A$ and for $\alpha \in \C$,
	
	\begin{enumerate}
		\item $a^{\ast\ast} = (a^\ast)^\ast = a$,
		\item $(\alpha a+b)^\ast = \overline{\alpha} a^\ast + b^\ast$,
		\item $(ab)^\ast = b^\ast a^\ast$.
	\end{enumerate}
	The element $a^\ast$ is referred to as the \emph{adjoint} of $a$. 		\\
	A \emph{$C^\ast$-algebra} is a Banach algebra $A$ with involution $a \mapsto a^\ast$ making it a 
	$\ast$-algebra, with the condition that
	\begin{align*}
		\|a ^\ast a\| = \|a\|^2 \mbox{ for all } a \in A.
	\end{align*}
	This condition is known as the \emph{$C^\ast$-axiom}.\\
	A subalgebra $B$ of a $C^\ast$-algebra $A$ is a \emph{$C^\ast$-subalgebra} if it is closed under the 
	adjoint and complete with respect to the norm; equivalently, if $B$ is itself a $C^\ast$-algebra.
\end{defn}

Given a subset $S$ of $A$, let $C^\ast(S)$ denote the \emph{$C^\ast$-subalgebra
of $A$ generated by $S$}, which is the smallest $C^\ast$-subalgebra of $A$ containing $S$.
\todo[inline]{quotient algebra}%%%

An element $a\in A$ of a $C^\ast$-algebra is called
\begin{itemize}
	\item	\emph{normal} if $a^\ast a=aa^\ast$;
	\item 	\emph{self-adjoint} if $a^\ast=a$;
	\item 	\emph{unitary} if $aa^\ast=a^\ast a = \1$.
\end{itemize}
The set of self-adjoint elements $A_{sa}$ forms a vector space over $\R$.
For $a$ in $A$, let $h=\tfrac{1}{2}(a+a^\ast)$ and $k=\tfrac{i}{2}(a-a^\ast)$. Then $h$ and $k$ are 
self-adjoint and we can write $a=h+ik$ -- call $h$ and $k$ the \emph{real} and \emph{imaginary} part of
$a$, respectively.



\section{Unitization}
If a $C^\ast$-algebra $A$ contains an identity element $\1$ such that 
$a \1 = a = \1  a$ for all $a \in A$, call $\1$ the \emph{unit} 
in $A$, and $A$ is then a \emph{unital} $C^\ast$-algebra.

\begin{prop*}[{\cite[I.1.3]{davidson96}}]
	Any non-unital $C^\ast$-algebra $A$ can be isometrically embedded in a unital $C^\ast$-algebra 
	$\tilde{A}$.
\end{prop*}
\begin{proof}
	Let $\tilde{A} = A \oplus \C$ with pointwise addition, and define
	\begin{align*}
		(a,\lambda) (b,\mu) &= (ab+\lambda b + \mu a, \lambda \mu),				\\
		(a,\lambda)^\ast &= (a^\ast,\overline{\lambda}).
	\end{align*}
	Then $\tilde{A}$ is a $\ast$-algebra. 
	Consider $(a,\lambda)$ as an operator acting on $A$ via $b\mapsto ab+\lambda b$ for $b$ in $A$. It 
	is easy to see that this operator is linear, and 
	\begin{align*}
				\sup_{\|b\|=1}\|ab+\lambda b\| 
		&\leq 	\sup_{\|b\|=1} (\|ab\|+ \lambda\|b\|)							\\
		&=		\sup_{\|b\|=1} \|ab\|+ \lambda < \infty
	\end{align*}
	so that the operator $(a,\lambda)$ is bounded. It follows that
	\begin{align*}
		\|(a,\lambda)\| = \sup_{\|b\|=1}\|ab+\lambda b\|
	\end{align*}
	is a norm on $\tilde{A}$, and so $\tilde{A}$ is a Banach $\ast$-algebra, with unit $(0,1)$.
	The embedding $a\mapsto(a,\lambda)$ is isometric because 
	\[
		\|a\| = \|a\cdot\frac{a}{\|a\|}\| \leq \|(a,0)\| \leq \sup_{\|b\|=1}{\|ab\|} \leq \|a\|.
	\]
	It remains to verify the $C^\ast$-axiom:
	\begin{align*}
				\|(a,\lambda)\|^2 
		&=		\sup_{\|b\|=1}{\|ab+\lambda b\|^2}								\\
		&=		\sup_{\|b\|=1}{\|b^\ast a^\ast ab 
								+\lambda b^\ast a^\ast b
								+\overline{\lambda}b^\ast a b
								+|\lambda|^2 b^\ast b}							\\
		&\leq	\sup_{\|b\|=1}{\|a^\ast ab 
								+\lambda a^\ast b
								+\overline{\lambda}a b
								+|\lambda|^2 b\|}								\\
		&=		\|(a^\ast a + \lambda a^\ast +\overline{\lambda}a,|\lambda|^2)	\\
		&= 		\|(a,\lambda)^\ast(a,\lambda)\|									\\
		&\leq	\|(a,\lambda)^\ast\| \|(a,\lambda)\|.
	\end{align*}
	By symmetry of $\ast$, $\|(a,\lambda)^\ast\| = \|(a,\lambda)\|$. 
	Hence, the above inequality becomes equality and we have that
	\begin{align*}
		\|(a,\lambda)^\ast(a,\lambda)\| = \|(a,\lambda)\|^2. 
	\end{align*}
\end{proof}

In light of this result, we take all $C^\ast$-algebras from here to be unital unless
specified otherwise.
For the results we will consider, we can simply consider the unital case. 
However, there are important circumstances in advanced theory in which one 
needs to relax the unital condition.


\section{$\CX$ - an example.}
Given a compact Hausdorff space $X$, let $\CX$ be the algebra of continuous functions
$f:X\to\C$, with addition and multiplication defined pointwise. Define $\|\cdot\|$ on $\CX$ by
\[
	\|f\|= \sup_{x\in X}{|f(x)|},
\]
and let $\overline f$ be the function $\overline f(x) = \overline{f(x)}$ for $x$ in $X$.
We will demonstrate that $\CX$ is an Abelian $C^\ast$-algebra and use it to prototype the theory
we will build up in later sections. Later, in Section \ref{section:abelian}, we will show that $\CX$ is 
essentially \emph{the} Abelian $C^\ast$-algebra.

\begin{claim}
	$\CX$ is an Abelian $C^\ast$-algebra.
\end{claim}



\todo [inline]{example of $C(X)$? could define states and stuff preemptively} %%



\section{The Spectrum}
\todo[inline]{write spectral theory section.} %%
We will just state the necessary definitions and results here, and defer proofs to sources.

\begin{defn}
	Given an element $a$ of a Banach algebra $A$, define its spectrum $\sigma_A(a)$:
	\[
		\sigma_A(a)=\{\lambda\in \C ~|~ a-\lambda\1 \mbox{ is not invertible in }A\},
	\]
	where $a$ is \emph{invertible in $A$} if there exists an element $b$ such that $ab=\1=ba$. 
	Say that $\lambda\in\spec a$ is a \emph{spectral value} of $a$ in $A$. If it is clear from context 
	where we are taking the spectrum, then we write $\spec{a}$ for $\sigma_A(a)$. 
\end{defn}

\begin{thm*}[{\cite[3.2.3]{kadison83}}]
	If $a$ is an element of a Banach algebra $A$ then $\spec a$ is a non-empty closed subset of the
	closed disk in $\C$ with center $0$ and radius $\|a\|$.
\end{thm*}

The \emph{spectral radius}, $r(a)$, of $a$ is the radius of the smallest disk in $\C$ containing 
$\spec a$; that is, 
\begin{align*}
	r(a)=\sup_{\lambda\in\spec a}{|\lambda|}.
\end{align*}
By the above theorem, $r(a)\leq\|a\|$. 

\begin{thm*}[Spectral radius formula {\cite[3.3.3]{kadison83}}]
	The spectral radius of an element $a$ of a Banach algebra is given by the formula
	\begin{align*}
		r(a)= \lim_{n\to\infty} \|a^n\|^{1/n}.
	\end{align*}
\end{thm*}

\begin{lemma}[{\cite[4.1.1(i)]{kadison83}}]
	If $a$ is a normal element of in a $C^\ast$-algebra, then $r(a)=\|a\|$.
\end{lemma}


\begin{thm*}[{\cite[4.1.5]{kadison83}}]
	If $B$ is a $C^\ast$-subalgebra of a $C^\ast$-algebra $A$, then $\sigma_B(b) =\sigma_A(b)$ for all 
	$b$ in $B$.
\end{thm*}
As an aside, this theorem allows us to make the convention that the spectrum of an element $a$ in a 
non-unital $C^\ast$-algebra $A$ is defined to be the spectrum of $a$ in the unitization of $A$.




With commuting, self-adjoint elements $a,b\in A_{sa}$, we have $(ab)^\ast = b^\ast a^\ast = ba =
ab$, so $ab$ is self-adjoint. Since each of $a,b,ab$ have the same spectrum in $A$ as in the Abelian 
$C^\ast$-subalgebra $C^\ast(\1,a,b)$, we have
\begin{align*}
	\spec{ab} \subseteq \spec{a}\spec{b}.
\end{align*}
\todo{what is this for again?}


\section{Further Definitions}

An element $a$ of a $C^\ast$-algebra $A$ is \emph{positive} if it is self-adjoint and $\spec a
\subseteq \R^+$. 
Denote the set of positive elements in $A$ by $A^+$. Then there is a partial order $\leq$ on 
$A_{sa}$ defined by 
\begin{align*}
	a\leq b ~~\iff~~ b-a \in A^+.
\end{align*}
The set of positive elements form a \emph{positive cone} in $A_{sa}$, which means that 
\begin{enumerate}
	\item 	$a\in A^+$ and $-a\in A^+$ implies that $a=0$,
	\item	$\alpha a+b\in A^+$ for all $a,b \in A^+$ and $\alpha\in \R^+$.
\end{enumerate}
The unit $\1$ is positive (with $\spec \1 = \{1\}$), and for any $a\in A_{sa}$ we have $-\|a\|\1\leq a\leq\|a\|\1$.
\todo{justify (4.2.3(ii))} %%
	
\begin{defn}
	Given Banach $\ast$-algebras $A$ and $B$, a map $\phi:A\to B$ is a \emph{$\ast$-homomorphism} 
	if it is an algebra homomorphism for which $\phi(a^\ast) = \phi(a)^\ast$ for all $a \in A$.
	If a $\ast$-homomorphism is one-to-one, call it a \emph{$\ast$-isomorphism}.
\end{defn}

\begin{prop}[{\cite[4.1.8]{kadison83}}] \label{prop:homo}
	Suppose $A$ and $B$ are $C^\ast$-algebras and $\phi:A\to B$ is a $\ast$-homomorphism. Then
	$\|\phi(a)\| \leq \|a\|$ for all $a \in A$. If $\phi$ is a $\ast$-isomorphism, then
	$\|\phi(a)\| = \|a\|$ for all $a \in A$.
\end{prop}
\begin{proof}
	
\end{proof}\todo[inline]{prove}%%


\begin{defn}
	A linear functional $\rho$ on a $C^\ast$-algebra $A$ is \emph{positive} if $\rho(a)\geq 0$ for all 
	$a\in A^+$. A \emph{state} on $A$ is a positive linear functional $\rho$ such that $\rho(\1)=1$ and
	$\rho (a) \geq 0$  for all positive elements $a \in A^+$. 
	Denote by $\S(A)$ the set of all states on $A$. \todo{states are real on self-adjoint elements}%%
\end{defn}

We can extend the partial order notation to linear functionals. Given linear functionals $\rho,\tau$ on $A$, define $\leq$ by
\begin{align*}
	\rho \leq \tau ~~\iff~~ \tau - \rho \mbox{ is positive}.
\end{align*}

\todo[inline]{geometric interpretation - for $A=\C$, self adjoint elements are real numbers and the
 positive cone is $\R^+$. } %%

\begin{prop}[{\cite[4.3.2]{kadison83}}]
	A linear functional $\rho$ on a $C^\ast$-algebra $A$ is positive if and only if $\rho$ is bounded
	and $\|\rho\| = \rho(\1)$.
\end{prop}
\begin{proof} \todo{rework; explain further.}%%
	Suppose that $\rho$ is positive. With $a$ in $A$, let $\alpha$ be a scalar of modulus $1$ 
	\todo{why does this exist?} %%
	such that $\alpha\rho(a)\geq 0$, and let $h$ be the real part of $a$.
	Since $h$ is self adjoint, (by 4.2.3(ii)) we have $h \leq \|h\|\1 \leq \|a\|\1$. Thus, $\|a\|\1-h$
	is positive and 
	\[
		\rho(\|a\|\1-h) = \|a\|\rho(\1)-\rho(h) \geq 0.
	\]
	Therefore,
	\begin{align*}
		|\rho(a)| = \rho(\alpha a) = \overline{\rho(\alpha a)} = \rho(\overline{\alpha}a^\ast) = 
					\rho(\tfrac{1}{2}(\alpha a +\overline{\alpha}a^\ast)) =
					\rho(h) \leq \rho(\1) \|a\|,
	\end{align*}
	so $\rho$ is bounded and $\|\rho\| \leq \rho(\1)$. We also have $\|\rho\| = \sup\{\rho(a)~|~\|a\|=
	1\} \geq \rho(\1)$, so $\|\rho\|=\rho(\1)$.
	
	Conversely, suppose $\rho$ is bounded and $\|\rho\| = \rho(\1)$ -- we can assume without loss that 
	$\rho(\1)=1$. With $a$ a positive element of $A$, let $\rho(a)=\alpha+i\beta$ for real $\alpha
	,\beta$. Then $\rho$ is positive if and only if $\alpha \geq 0$ and $\beta =0$.
	For small positive $s$, 
	\begin{align*}
		\spec{\1-sa} = \{1-st ~|~ t\in \spec{a} \subseteq \R^+\} \subseteq [0,1],
	\end{align*}
	so $\|\1-sa\| = r(\1-sa) \leq 1$. Hence
	\begin{align*}
		1-s\alpha \leq |1-s(\alpha+i\beta)| = |\rho(\1 - sa)| \leq 1,
	\end{align*}
	so $\alpha\geq0$. With $b_n$ in $A$ defined by $b_n= a+(in\beta-\alpha)\1$ for each positive
	integer $n$,
	\begin{align*}
				\|b_n\|^2 = \|b_n^\ast b_n\| 
		&= 		\|(a-\alpha\1)^2 + n^2\beta^2\1\| 									\\
		&\leq 	\|a-\alpha\1\|^2 +n^2\beta^2.
	\end{align*}
	Hence for all positive integers $n$, we have
	\begin{align*}
				(n^2+2n+1)\beta 
		&=	 	|\rho(b_n)|^2  														\\
		&\leq 	\|a-\beta\1\|^2 +n^2\beta^2,
	\end{align*}
	so that $\beta=0$. We conclude that $\rho$ is positive.
\end{proof}

\begin{prop*}[Cauchy-Schwarz inequality, {\cite[4.3.1]{kadison83}}]
	If $\rho$ is a positive linear functional on $C^\ast$-algebra $A$, then for all $a$ and $b$ in $A$,
	\begin{align*}
		|\rho(b^\ast a)|^2 \leq \rho(a^\ast a)\rho(b^\ast b).
	\end{align*}
\end{prop*}
\todo{prove?}


\begin{lemma}[{\cite[4.3.3]{kadison83}}]
	Let $A$ be a $C^\ast$-algebra. For any $a$ in $A$ and $\alpha\in\spec{a}$, there exists a state
	$\rho$ on $A$ such that $\rho(a)=\alpha$.
\end{lemma}
\begin{proof}
	For all complex numbers $\beta$ and $\gamma$, $\alpha\beta+\gamma$ is a spectral value for the 
	element $\beta a+\gamma\1$ of $A$, so 
	\[
		|\alpha\beta+\gamma|\leq r(\beta a+\gamma\1)=\|\beta a+\gamma\1\|.
	\] 
	Hence the equation $\rho_0(s)=\alpha\beta+\gamma$ defines a bounded linear functional
	$\rho_0$ on the linear subspace $B= \{\beta a+\gamma\1 ~|~ \beta,\gamma\in\C\}$ of $A$, with 
	$\rho_0(a) =\alpha$ and $\rho_0(\1)=1=\|\rho_0\|$. By the Hahn-Banach theorem, 
	$\rho_0$ extends to a bounded linear functional $\rho$ on $A$, with $\|\rho\|=1$, such that 
	$\rho=\rho_0$ on the subspace $B$. In particular, $\rho(\1)=1=\|\rho\|$ so $\rho$ is positive by 
	the previous result, and $\rho(a)=\alpha$.
\end{proof}

\begin{lemma}[{\cite[4.3.4,(i)]{kadison83}}]\label{lemma:state1}
	Let $A$ be a $C^\ast$-algebra. If $\rho(a)=0$ for all states $\rho$ on $A$, then $a=0$.
\end{lemma}
\begin{proof}
	Suppose first that $a$ is self-adjoint, and $\rho(a)=0$ for all states $\rho$. By the previous
	result, $\spec{a}=\{0\}$, so $\|a\|=r(a)=0$. Hence $a=0$.			\\
	Now write $a=h+ik$, for $h$ and $k$ the real and imaginary part of $a$ respectively. Then
	\begin{align*}
		0=\rho(a)=\rho(h)+i\rho(k),
	\end{align*}
	and as $h,k$ are self-adjoint, $\rho(h)$ and $\rho(k)$ are real and we must have 
	$\rho(h)=0=\rho(k)$. By previous statement, $h=0=k$, and we conclude that $a=0$.
\end{proof}

\begin{defn}
	An extreme point of $\S(A)$ is called a \emph{pure} state on $A$, and the set of pure 
	states on $A$ is denoted by $\P{A}$.
\end{defn}
The state space $\S(A)$ is then the weak$^\ast$-closure of the set of convex linear combinations of pure 
states. 
It is a simple exercise, using the fact that $A^+\subseteq A_{sa}$, to verify that a linear 
functional $\rho$ is a pure state on $A$ if and only if the restriction $\rho|_{A_{sa}}$ is a pure
state on $A_{sa}$.
Every pure state on $A_{sa}$ extends to a pure state on $A$ with the same norm, by the Hahn-Banach 
theorem. We will need the following few results on pure states later.

\begin{lemma}[{\cite[4.3.8,(i)]{kadison83}}]\label{lemma:pure1}
	Let $A$ be a $C^\ast$-algebra. If $\rho(a)=0$ for all \emph{pure} states $\rho$ on $A$, then $a=0$.
\end{lemma}
\begin{proof}
	Since every state is the limit of a sequence of linear combinations of pure states, if $\rho(a)=0$ 
	for all pure states $\rho$ on $A$, then $\rho(a)=0$ for \emph{all} states $\rho$ on $A$. The result 
	follows immediately from Lemma \ref{lemma:state1}.
\end{proof}

\begin{prop}[{\cite[3.4.6]{kadison83}}]\label{prop:pure2}
	A state $\rho$ on $A_{sa}$ is pure if and only if, for all positive linear functionals $\tau$ on
	 $A_{sa}$ such that $0\leq\tau\leq\rho$, we have $\tau =\lambda\rho$ for some $\lambda\in\R$. 
\end{prop}
\begin{proof}
	Suppose that $\tau =\lambda\rho$ for all $0\leq\tau\leq\rho$, and suppose we can write 
	$\rho=\alpha\rho_1+(1-\alpha)\rho_2$ for some $0\leq\alpha\leq1$ and some $\rho_1,\rho_2 \in
	\S(A_{sa})$. Then $0\leq\alpha\rho_1\leq\rho$, so $\alpha\rho_0 = \lambda\rho$. Then 
	$\rho_1(\1)=1=\rho(\1)$, so 
	$\alpha = \lambda$ so $\rho_0 =\rho$. Similarly, we can show that $\rho_2=\rho$, and so we
	 conclude that $\rho$ is a pure state.
	
	Conversely, suppose that $\rho$ is a pure state and $0\leq\tau\leq\rho$. Applying this to $\1$,
	we get $0\leq\tau(\1)\leq\rho(\1) = 1$. Let $\lambda = \tau(\1)$. We work case-by-case: \\
	If $\lambda=0$, then for any $a\in A_{sa}$, applying $\tau$ to $-\|a\|\1\leq a\leq\|a\|\1$ gives
	\begin{align*}
		0=-\|a\|\lambda=\tau(-\|a\|\1)\leq\tau(a)\leq\tau(\|a\|\1)=\|a\|\lambda=0,
	\end{align*}
	so $\tau=0=\lambda\rho$. \\
	A similar argument shows that $\lambda=1$ implies $\tau-\rho=0$ so that $\tau=\rho=\lambda\rho$.\\
	If $0\leq\lambda\leq 1$, we can write $\rho=\lambda\rho_1+(1-\lambda)\rho_2$ for 
	$\rho_1=\lambda^{-1}\tau$ and $\rho_2=(1-\lambda)^{-1}(\rho-\tau)$. $\rho$ is pure so 
	$\tau=\lambda\rho_1=\lambda\rho$.
	
\end{proof}
A linear functional $\rho$ on a $C^\ast$-algebra $A$ is \emph{multiplicative} if $\rho(ab)=\rho(a)
\rho(b)$ for all $a,b$ in $A$. The set of all multiplicative linear functionals on $A$ is called the 
\emph{maximal ideal space} of $A$, denoted $\M_A$. This name hints at the fact that the kernel of each 
of these functionals is a maximal ideal of $A$, and all maximal ideals of $A$ arise in this way. 
\cite[Theorem I.2.5]{davidson96}
\begin{prop}[{\cite[4.4.1]{kadison83}}]\label{prop:puremult}
	The set of pure states on an Abelian $C^\ast$-algebra $A$ is precisely the maximal ideal space of 
	$A$.
\end{prop}
\begin{proof}
	Suppose $\rho$ is a pure state on $A$. To show that $\rho(ab)=\rho(a)\rho(b)$ for $a,b,\in A$,
	we restrict attention to the case where $0\leq b\leq\1$. Linearity gives us the general case.
	\todo{show} %%
	In this case, for $h\in A^+$ we have that $0\leq hb\leq h$, so $0\leq\rho(hb)\leq\rho(h)$.
	Hence $\rho_b(a)=\rho(ab)$ for $a\in A$ defines a positive linear functional on $A$ with 
	$\rho_b\leq\rho$. The restriction $\rho|_{A_{sa}}$ is a pure state on $A_{sa}$ and 
	$\rho_b|_{A_{sa}} \leq \rho|_{A_{sa}}$, and it follows from Proposition \ref{prop:pure2} that 
	$\rho_b|_{A_{sa}} = \alpha \rho|_{A_{sa}}$ for some $\alpha \in \R^+$.
	Hence $\rho_b = \alpha\rho$  and so for $a\in A$:
	\begin{align*}
		\rho(ab) = \rho_b(a) = \alpha\rho(a) = 
							\alpha\rho(\1)\rho(a) = \rho_b(\1)\rho(a) = \rho(b)\rho(a)
	\end{align*}
	
	Conversely, suppose $\rho$ is a multiplicative linear functional. \todo[inline]{show a state.} %% 
	Suppose we can write $\rho=\alpha\rho_1+\beta\rho_2$ for states $\rho_1,\rho_2$ on $A$ and 
	$\alpha,\beta >0$ such that $\alpha+\beta=1$. For $c\in A_{sa}$, by the Cauchy-Schwarz inequality
	 we have for $j=1,2$:
	\begin{align*}
		\big(\rho_j(c)\big)^2 = \big(\rho_j(\1 c)\big)^2 \leq \rho_j(\1)\rho(c^2)=\rho(c^2).
	\end{align*}
	Then:
	\begin{align*}
				0
		&=		\rho(c^2)-\rho(c)^2 											\\
		&=		\alpha\rho_1(c^2)+\beta\rho_2(c^2) 
						- \big(\alpha\rho_1(c)+\beta\rho_2(c)\big)^2			\\
		&\geq	\alpha(\alpha+\beta)\rho_1(c)^2 
						+ \beta(\alpha+\beta)\rho_2(c)^2
						- \big(\alpha\rho_1(c)+\beta\rho_2(c)\big)^2			\\
		&=		\alpha\beta\big(\rho_1(c) - \rho_2(c)\big)^2.
	\end{align*}
	Hence $\rho_1(c)=\rho_2(c)$, for all $c\in A_{sa}$, so $\rho_1=\rho_2$ and we conclude
	that $\rho$ is a pure state. 
\end{proof}




\todo[inline]{discuss weak* topology on P(S)}  %%


\section{$\B(\H)$ - an example.}
This section concerns the fundamental example of a $C^\ast$-algebra - the set $\B(\H)$ of bounded 
linear operators on a Hilbert space $\H$.  
Here we will demonstrate that $\B(\H)$ is a $C^\ast$-algebra and give some basic results.

\begin{claim} $\B(\H)$ is a $C^\ast$-algebra with the operator norm 
\[
	\|T\|= \sup_{\|x\|=1}{\|Tx\|}
\]
and involution taking $T$ to its adjoint map $T^\ast$. 
The identity map $I:x\mapsto x$ is a unit for $\BH$
\end{claim}
\begin{proof}
	$\|\cdot\|$ is a norm on $\BH$. Let $\{T_n\}_{n\in\N}$ be a Cauchy sequence 
	in $\BH$. Then for any positive $\epsilon$, there is a positive integer $N$ such that 
	\begin{align*}
		\|T_m-T_n\| < \epsilon \mbox{ for all } m,n \geq N.
	\end{align*}
	Applying $T_m-T_n$ to $x \in \H$, we have 
	\begin{align}\label{eqn:cauchyTn}
		\|T_mx-T_nx\| \leq \|T_m-T_n\| \|x\| < \epsilon \|x\|,
	\end{align}
	so $\{T_nx\}_{n\in\N}$ is a Cauchy sequence in $\H$, converging to an element in $\H$.
	Define a linear operator $T:\H \to \H$ by 
	\begin{align*}
		Tx= \lim_{n\to\infty}{T_nx} \mbox{ for } x \in \H.
	\end{align*}
	Taking limits as $m$ tends to infinity in equation \eqref{eqn:cauchyTn}, we obtain
	\begin{align*}
		\|Tx-T_nx\| < \epsilon \|x\| \mbox{ for all }n \geq N,
	\end{align*}
	and so we have that $T-T_n$ (and hence $T=(T-T_n)+T_n$) is a bounded operator and  
	\begin{align*}
		\|T-T_n\| <\epsilon \mbox{ for all }n \geq N.
	\end{align*}
	We conclude that $T_n \to T$, and so $\BH$ is complete.
	
	Since boundedness is equivalent to continuity on $\H$, given $S,T\in\BH$, the operator 
	$ST:\H \to \H; x \mapsto (S\circ T)(x)$ is bounded on $\H$.
	Given $x\in\H$ and $\lambda\in\C$, 
	\begin{align*}
			((\lambda S)T)(x)
		&=	((\lambda S)\circ T)(x)												\\
		&=	\lambda S(Tx)														\\
		&=	\lambda (S\circ T)(x)												\\
		&=	\lambda ST(x),
	\end{align*} 
	so that $(\lambda S)T = \lambda ST$ in $\BH$. We have 
	\begin{align*}
				\|ST\|
		&=		\sup_{\|x\|=1}{\|STx\|} 										\\
		&=		\sup_{\|x\|=1}{\|S(Tx)\|} 										\\
		&\leq	\|S\| \sup_{\|x\|=1}{\|Tx\|} 									\\
		&=		\|S\| \|T\|.
	\end{align*}
	We conclude that $\BH$ is a Banach algebra.
	
	To see that $^\ast$ is an involution, use the fact that the adjoint 
	operator is unique for each operator and the following equalities.
	\begin{enumerate}
	\Item	\begin{align*}
				\langle (\alpha T+S)^\ast x,y\rangle 
			&=	\langle x, \alpha T+S y \rangle									\\
			&=	\overline{\alpha}\langle x,Ty\rangle + \langle x, Sy \rangle	\\
			&=	\overline{\alpha}\langle T^\ast x,y\rangle +
										 \langle S^\ast x, y \rangle			\\
			&=	\langle(\overline{\alpha} T^\ast + S^\ast) x, y \rangle.
			\end{align*}
	\Item 	\begin{align*}
				\langle (T^\ast)^\ast x,y\rangle 
			&=	\langle x, T^\ast y \rangle										\\
			&=	\overline{\langle T^\ast y,x \rangle}							\\
			&=	\overline{\langle y, Tx \rangle}								\\
			&=	\langle Tx,y\rangle.
			\end{align*}
	\Item	\begin{align*}
				\langle (ST)^\ast x,y \rangle
			&=	\langle x, STy \rangle											\\
			&=	\langle	S^\ast x,Ty \rangle										\\
			&=	\langle T^\ast S^\ast x,y \rangle.
		\end{align*}
	\end{enumerate}
	It remains to demonstrate the $C^\ast$-axiom on $\BH$. For all $x\in\H$, we have
	\begin{align*}
		\|Tx\|^2 = \langle Tx,Tx\rangle = \langle T^\ast Tx,x\rangle \leq \|T^\ast T\| \|x\|^2,
	\end{align*}
	so that
	\begin{align*}
		\|T\|^2 \leq \|T^\ast T\| \leq \|T^\ast\| \|T\| = \|T\|^2.
	\end{align*}
	It is clear that $I$ is a unit.
	Hence, the claim.
\end{proof} 




\todo [inline]{example of $\B(\H)$ at end of section to retro-motivate notation.
				discuss nomenclature (state etc) coming from QM}%%




\chapter{Representations of $C^\ast$-algebras}
\section{Abelian $C^\ast$-algebras}\label{section:abelian}

\todo[inline]{rewrite}%%

Let $A$ be an Abelian $C^\ast$-algebra. For $a$ in $A$, define a complex-valued function $\hat{a}$ on
$\P{A}$ by $\hat{a}(\rho)= \rho(a)$. The weak$^\ast$-topology is the weakest topology on 
$\P{A}$ for which all of the maps $\hat{a}$ are continuous, so that $\hat{a} \in C(\P{A})$ for all 
$a\in A$. The map
\[
	\Gamma:A \to C(\P{A}) : a\mapsto \hat{a}
\]
is called the \emph{Gelfand transform} of $A$ \cite{davidson96}.
For $a,b \in A$, $\alpha,\beta\in\C$ and $\rho\in\P{A}$:
\begin{align*}
	\widehat{(\alpha a+\beta b)}(\rho) = \rho(\alpha a+\beta b) 
	&=	\alpha\rho(a) +\beta\rho(b) = \alpha\hat{a}(\rho)+ \beta\hat{b}(\rho), 	\\
	\widehat{a^\ast}(\rho) = \rho(a^\ast) &= \overline{\rho(a)} = \overline{\hat{a}(\rho)}.
\end{align*}
Since by Proposition \ref{prop:puremult},
\begin{align*}
	\P{A} = \{\rho \in \S(A) ~|~ \rho(ab) = \rho(a)\rho(b)\},
\end{align*}
we have that 
\begin{align*}
	\widehat{(ab)} (\rho)= \rho(ab) = \rho(a)\rho(b) = \hat{a}(\rho)\hat{b}(\rho).
\end{align*}
Hence the Gelfand transform is a $\ast$-homomorphism. The following theorem gives us that it is
in fact an isometric $\ast$-isomorphism.

\begin{thm}[{Gelfand-Naimark, commutative \cite[4.4.3]{kadison83}}] 
	Every Abelian $C^\ast$-algebra $A$ is $\ast$-isomorphic to $\CX$, the algebra of continuous
	 functions on a compact Hausdorff space $X$.
\end{thm}
\begin{proof}
	
	
\end{proof}\todo[inline]{prove this.}%%


The previous result generalises to not-necessarily-unital Abelian $C^\ast$-algebras as follows:
\begin{thm*}[{\cite[I.2.7]{davidson96}}]
	Every Abelian $C^\ast$-algebra $A$ is $\ast$-isomorphic to $C_0(X)$, the algebra of 
	continuous functions on a locally compact Hausdorff space $X$ which vanish at infinity.
\end{thm*}
\begin{proof}
	\todo[inline]{citation needed}%%
\end{proof}
The unitization of this algebra then corresponds to the one-point compactificaton of $X$.
	
\section{The Gelfand-Naimark Theorem}

\begin{defn}
	Given a $C^\ast$-algebra $A$, a \emph{representation of $A$ on a Hilbert space $\H$} 
	is a $\ast$-homomorphism $\phi: A \to \BH$. 
	A $\ast$-isomorphic representation is called \emph{faithful}.
	If there exists an element $x\in\H$ such that the set $\{\phi(a) ~|~ a\in A\}$ is dense in $\H$, say 
	that $\phi$ is a \emph{cyclic} representation, with \emph{cyclic vector} $x$.
\end{defn}	
The construction used in the proof of the following theorem is known as the Gelfand-Naimark-Segal (GNS)
construction.
\begin{thm}[{\cite[4.5.2]{kadison83}}]
	If $\rho$ is a state on a $C^\ast$-algebra $A$, then there exists a cyclic representation 
	$\pi_\rho$ of $A$ on a Hilbert space ${H}_\rho$, with unit cyclic vector $x_\rho$, such that 
	\[ 
		\rho(a)= \langle \pi_\rho (a) x_\rho, x_\rho \rangle, ~~~~ \forall a \in A.
	\]
\end{thm}
\begin{proof}
	We will construct from $\rho$ the space $\Hr$,  representation $\pi_\rho$, 
	and vector $x_\rho$, and demonstrate the required properties.
	
	Consider the \emph{left kernel} of $\rho$:
	\[
		L_\rho = \{t \in A ~|~ \rho (t ^\ast t) = 0 \}.
	\]	
	For $a,b \in A$, define $\langle a , b \rangle_0 = \rho(b^\ast a)$, then $\langle \cdot, \cdot \rangle_0$ satisfies
	\begin{enumerate}
	  \item Linearity in 1st argument: for $a,b\in A$, $\alpha, \beta \in \mathbb{C}$:
		\begin{align*}
		   \langle \alpha a + \beta b, c \rangle_0 
		&= \rho (c^\ast(\alpha a + \beta b)   								\\
		&= \rho (\alpha c^\ast a + \beta c^\ast b)  						\\
		&= \alpha \rho (c^\ast a) + \beta \rho (c^\ast b)					\\
		&= \alpha \langle a , c \rangle_0 + \beta \langle b, c \rangle_0.
		\end{align*}
	  \item Conjugate symmetric: for $a,b \in A$:
	  	\begin{align*}
	  	   \langle b,a \rangle _0 
	  	&= \rho (a^\ast b)													\\
	  	&= \rho ((b^\ast a)^\ast)											\\
	  	&= \overline{\rho (b^\ast a)}										\\
	  	&= \overline{\langle a,b \rangle _0 }.
	  	\end{align*}
	  \item Positive semi-definite: for $a \in A$:
	  	\begin{align*}
	  	&=	\langle a,a\rangle_0 = \rho(a^\ast a) \geq 0,
	  	\end{align*}
	  	since $a^\ast a$ is positive for all $a$. \todo{show: 4.2.5/6}%%
	\end{enumerate}
	Note that $\langle \cdot, \cdot \rangle$ is not necessarily positive definite on 
	$A$ -- the left kernel is exactly where this fails.
	
	To see that $L_\rho$ is a linear subspace of $A$, consider 
	\[
		L= \{t \in A  ~|~ \langle t,a \rangle _0 = 0, ~\forall a \in A \}\subseteq L_\rho.
	\]
	For $t \in L_\rho$, by the Cauchy-Schwarz inequality we have 
	\[ 
		|\langle t,a \rangle_0|^2 \leq \langle t,t \rangle_0 \langle a,a\rangle_0,~~ \forall a \in A;
	\]
	that is,
	\[
		\langle t,a \rangle _0 = 0, ~~~ \forall a \in A,
	\]
	so $t\in L$ and $L_\rho =L$.
	Now, for $a,b \in L$, $\alpha \in \mathbb{C}$ and $c \in A$:
	\[
		\langle \alpha a + b, c \rangle _0 = \alpha \langle a,c \rangle _0 + 
						\langle b,c\rangle _0 = 0,
	\]
	so $\alpha a +b \in L$; also, $\langle 0,c\rangle _0 = 0$ so $0 \in L$. 
	Hence, $L_\rho  $ is a linear subspace of $A$.
	
	For $s \in A$, $t\in L_\rho$, by the Cauchy Schwarz inequality we have 
	\begin{align*}
		|\rho (s^\ast t) |^2 
		&= 		|\langle t,s\rangle_0 |^2 									\\
		&\leq 	\langle t,t\rangle_0 \cdot \langle s,s\rangle_0  			\\
		&= 		\rho (t^\ast t) \cdot \rho (s^\ast s)						\\
		&=		0,
	\end{align*}
	so $\rho (s^\ast t) = 0$. Letting $s = a^\ast a t$ for $a \in A$, then
	\begin{align*}
		\rho ((at)^\ast at) 
		&= 		\rho (at^\ast a^\ast at) 									\\
		&= 		\rho ((a^\ast at)^\ast t) 									\\
		&= 		\rho (s^\ast t) 											\\
		&=		0,
	\end{align*}
	so that $at \in L_\rho$, for all $a \in A$ and $t \in L_\rho$; 
	we conclude that $L_\rho$ is a left ideal in $A$. \todo{why do we need closure?} Closure  of $L_\rho$ follows from the fact that 
	it is the preimage in $A$ of $\{0\}$ under the continuous map $t \mapsto \rho (t^\ast t)$. 
	
	Consider now the quotient space $V_\rho = A / L_\rho$, with $\langle \cdot,\cdot\rangle$ defined by 
	\[
		\langle a+L_\rho,b+L_\rho \rangle = \langle a,b\rangle_0, ~~~~ 
		\mbox{for}~~a+L_\rho,b+L_\rho \in V_\rho.
	\]
	It follows from properties $i),~ii)$ and $iii)$ of $\langle \cdot,\cdot  \rangle _0$ that
	$\langle \cdot,\cdot  \rangle$ is an inner product on $V_\rho$ -- with
	\begin{align*}
				\langle a + L^\rho, a + L^\rho \rangle = 0
		&\iff 	\langle a,a\rangle_0=0										\\
		&\iff 	a \in L_\rho												\\
		&\iff 	a+L_\rho = 0+L_\rho
	\end{align*}
	giving positive definiteness.
	The completion of $V_\rho$ with respect to the induced norm $\|\cdot\|$ is a Hilbert space -
	this is the Hilbert space $\Hr$ we're looking for.
	
	Now we fix $a \in A$, and consider the map 
	\[
		\pi_a : V_\rho \to V_\rho; b+L_\rho \mapsto ab+L_\rho.
	\]
	Let $b_1, b_2 \in A$ be such that $b_1+L_\rho = b_2+L_\rho$. Then:
	\begin{align*}
		&\implies	b_1-b_2 \in L_\rho										\\	
		&\implies	a(b_1-b_2) \in L_\rho									\\	
		&\implies	ab_1-ab_2 \in L_\rho									\\	
		&\implies	ab_1 + L_\rho = ab_2+L_\rho								\\	
		&\implies	\pi_a(b_1+L_\rho) = \pi_a(b_2+L_\rho).
	\end{align*}
	Hence $\pi_a$ defines a linear operator on $V_\rho$.
	For $b+L_\rho \in V_\rho$:
	\begin{align*}
				\|a\|^2 \cdot \|b+L_\rho\| - \| \pi_a(b+L_\rho) \|
		&=		\|a\|^2 \cdot \|b+L_\rho\| - \| ab+L_\rho \|				\\
		&=		\|a\|^2 \cdot \langle b+L_\rho,b+L_\rho \rangle - 
									\langle ab+L_\rho,ab+L_\rho \rangle		\\
		&=		\|a\|^2 \cdot \rho(b^\ast b) - \rho ((ab)^\ast ab)			\\
		&= 		\rho (\|a\|^2 b^\ast b-b^\ast a^\ast ab)					\\
		&=		\rho (b^\ast (\|a\|^{2} \mathbbm{1} - a ^\ast a)b)			\\
		&\geq 	0.
	\end{align*}
	Thus $\pi_a$ is a bounded operator on $\Vr$, with $\|\pi_a\| \leq \|a\|$. By continuity,
	$\pi_a$ extends to a bounded operator $\pi_\rho(a):\Hr \to \Hr$,
	such that \[ \pi_\rho(a)(v) = \pi_a(v) \] for $v\in \Vr$. 
	Then $\pi_\rho(a) \in \B(\Hr)$ for each $a \in A$, so $\pi_\rho$ defines a map 
	$A \to \B(\Hr)$ such that $ a \mapsto \pi_\rho(a)$. This will be our representation.
	
	Now, for $a,b \in A$, $c+ L_\rho \in \Vr$ and $\alpha \in \mathbb{C}$:
	\begin{align*}
				\pi_{\alpha a+b}(c+\Lr)
		&=		(\alpha a+b) (c+\Lr)										\\
		&=		(\alpha ac +\Lr) + (bc+\Lr)									\\
		&=		\alpha \pi_a (c+\Lr) + \pi_b(c+\Lr),
	\end{align*}
	so that $\pi_{\alpha a + b} = \alpha \pi_a +\pi_b$ on $\Vr$.\\
	For $a,b \in A$ and $c+ L_\rho \in \Vr$:
	\begin{align*}
		\pi_{ab}(c+\Lr)
		&=		abc+\Lr														\\
		&=		\pi_a (bc+\Lr)												\\
		&=		\pi_a (\pi_b (c+\Lr))										\\
		&=		(\pi_a \cdot \pi_b) (c+\Lr),
	\end{align*}
	so that $\pi_{a b} = \pi_a \cdot \pi_b$ on $\Vr$.\\
	For $a\in A$ and $b+ L_\rho,~c+ L_\rho \in \Vr$:
	\begin{align*}
				\langle b+\Lr, \pi_a^\ast (c+\Lr) \rangle 
		&=		\langle \pi_a (b+\Lr), c+\Lr \rangle						\\
		&=		\langle ab +Lr, c+\Lr \rangle								\\
		&=		\rho(c^\ast ab)												\\
		&=		\rho((a^\ast c)^\ast b)										\\
		&=		\langle b+\Lr, a^\ast c+\Lr \rangle							\\
		&=		\langle b+\Lr, \pi_{a^\ast}(c+\Lr),
	\end{align*}
	so that $ \pi_a^\ast = \pi_{a^\ast}$ on $\Vr$.\\
	$\Vr \subset \Hr$ is a dense subset, so the three properties above hold on 
	$\Hr$ by continuity of $\pi_\rho$. 
	Hence, $\pi_\rho: A \to \B(\Hr)$ is a representation of $A$.
	As to the unit vector, consider $\xr = \1 + \Lr \in \Vr$. Then for $a \in A$,
	\begin{align*}
				\langle \pi_\rho(a)\xr ,\xr \rangle 
		&=		\langle \pi_a(\1+\Lr), \1+\Lr \rangle						\\
		&=		\langle a+\Lr \1+\Lr \rangle								\\
		&=		\rho(a);
	\end{align*}
	and in particular, $\langle \xr,\xr \rangle = \rho (\1) = 1$, so $\xr$ is a unit vector in $\Hr$.	
\end{proof}
\todo[inline]{example of construction on C(X)? short explanation of how B(H) and 
C(X) link together. can then talk about noncommutative topology!} %%



\begin{thm}[Gelfand-Naimark, {\cite[4.5.6]{kadison83}}]
	Every $C^\ast$-algebra has a faithful representation.
\end{thm}
\begin{proof}
	Let $A$ be a $C^\ast$-algebra. Suppose we have a collection $\{\phi_i ~|~i\in I\}$ of 
	representations of $A$ on Hilbert spaces $\{\H_i ~|~ i\in I\}$.
	For $a$ in $A$, we have $\|\phi_i(a)\|\leq \|a\|$ (by Proposition \ref{prop:homo}, as each $\phi_i$ 
	is a $\ast$-homomorphism), so we have a bounded operator $\oplus\phi_i(a)$ on $\oplus \H_i$ by 
	section \ref{section:hscons}. By the properties of $\oplus\phi_i(a)$ stated therein, the map 
	\[
		\phi:A\to \B(\oplus\H_i):a\mapsto\oplus\phi_i(a)
	\]
	is a $\ast$-homomorphism, and so is a representation of $A$ on $\oplus\H_i$. Call $\phi$ the 
	\emph{direct sum} of the collection $\{\phi_i ~|~i\in I\}$, denoted $\oplus\phi_i$.
	
	With $\S_0$ any collection of states on $A$ containing $\P{A}$, let $\phi$ be the direct sum of
	the collection $\{\pi_\rho ~|~ \rho\in\S_0\}$ of representations as constructed by the GNS
	construction. We will show that $\phi$ is a faithful representation.
	
	Given $a$ in $A$, if $\phi(a)=0$ then $\pi_\rho(a)=0$ for all pure states $\rho$ on $A$. But then,
	since $\rho(a)=\langle \pi_\rho (a) x_\rho, x_\rho \rangle$ by the GNS construction, we have 
	$\rho(a)=0$, and then by Lemma \ref{lemma:pure1}, $a=0$. Thus, $\phi$ is one-to-one and hence a 
	faithful representation of $A$.
\end{proof}




\todo[inline]{different reps given by different $\S_0$s.}
\todo[inline]{further topics: von neumann algebras (formal defn), K-theory, group C* algebras, amenable algebras, } %%
\todo[inline]{references!!!!}%%


\begin{thebibliography}{00}
\bibitem{blackadar06}
	Blackadar, B.,
	\emph{Operator Algebras: Theory of C*-Algebras and von Neumann Algebras.}
	Encyclopaedia of Mathematical Sciences, 122. Operator Algebras and Non-commutative Geometry, III. Springer-Verlag, Berlin (2006).
	
\bibitem{bornjordan25}
	Born, M. \& Jordan, P.,
	\emph{Zur Quantenmechanik.}
	Z. Physik (1925) 34: 858.
	
\bibitem{bornjordanheisenberg25}
	Born, M.; Heisenberg, W. \& Jordan, P.,
	\emph{Zur Quantenmechanik. II.}
	Z. Physik (1926) 35: 557.

\bibitem{davidson96}
	Davidson, K.R.,
	\emph{C*-Algebras by example.}
	Fields Institute Monographs, 6. American Mathematical Society, Providence, RI (1996).

\bibitem{dixmier77}
	Dixmier, J.,
	\emph{C*-algebras.}
	Holland Mathematical Library, Vol. 15. North-Holland Publishing Co., Amsterdam -- New York -- Oxford (1977).

\bibitem{dixmier81}
	Dixmier, J.,
	\emph{von Neumann algebras.}
	North-Holland Publishing Co., Amsterdam -- New York (1981).
		
\bibitem{elliott94}
	Elliott, G. A.,
	\emph{The classification problem for amenable C*-algebras.}
	Proceedings of the International Congress of Mathematicians, Vol. 1, 2 (Z\"{u}rich, 1994), pp.~922-–932, Birkhäuser, Basel (1995).

\bibitem{fell61}
	Fell, J. M. G.,
	\emph{The structure of algebras of operator fields. }
	Acta Math. 106, 1961, 233–280. 


\bibitem{gardella15}
	Gardella, E. E.,
	\emph{Compact group actions on C*-algebras: classification, non-classifiability, and crossed products and rigidity results for L$^p$-operator algebras. }
	Thesis (Ph.D.) -– University of Oregon (2015).

\bibitem{gelfand43}
	Gelfand, I. \& Neumark, M.,
	\emph{On the imbedding of normed rings into the ring of operators in Hilbert space.}
	Rec. Math. [Mat. Sbornik] N.S. 12(54) (1943), pp.~197–-213.

\bibitem{heisenberg25}
	Heisenberg, W.,
	\emph{{\"U}ber quantentheoretische Umdeutung kinematischer und mechanischer Beziehungen.}
	Z. Physik (1925) 33: 879. 
	
\bibitem{kadison83}
	Kadison, R. V. \& Ringrose, J. R.,
	\emph{Fundamentals of the theory of operator algebras: Vol. I. Elementary theory.}
	Pure and Applied Mathematics, 100. Academic Press, Inc. [Harcourt Brace Jovanovich, Publishers], New York (1983).

\bibitem{kadison86}	
	Kadison, R. V. \& Ringrose, J. R.,
	\emph{Fundamentals of the theory of operator algebras: Vol. II. Advanced theory.}
	Pure and Applied Mathematics, 100. Academic Press, Inc., Orlando, FL (1986).
	
\bibitem{lin01}
	Lin, H.,
	\emph{An introduction to the classification of amenable C*-algebras.}
	World Scientific Publishing Co., Inc., River Edge, NJ (2001).
\bibitem{lin08}
	Lin, H. \& Niu, Z.,
	\emph{Lifting KK-elements, asymptotic unitary equivalence and classification of simple C*-algebras.}
	Adv. Math. 219 (2008), no. 5, pp.~1729-–1769. 
	
\bibitem{lin11}
	Lin, H.,
	\emph{Asymptotic unitary equivalence and classification of simple amenable C*-algebras.}
	Invent. Math. 183 (2011), no. 2, pp.~385-–450. 

\bibitem{mackinnon77}
	MacKinnon, E.,
	\emph{Heisenberg, Models, and the Rise of Matrix Mechanics.}
	Hist. Stud. Phys. Sci., Vol. 8 (1977), pp.~137--188
	
\bibitem{vonneumann35}
	Murray, F. J. \& von Neumann, J.,
	\emph{On rings of operators.}
	Ann. of Math. (2) 37 (1936), no. 1, pp.~116–-229.

\bibitem{vonneumann37}
	Murray, F. J. \& von Neumann, J.,
	\emph{On rings of operators. II.}
	Trans. Amer. Math. Soc. 41 (1937), no. 2, pp.~208-–248. 

\bibitem{vonneumann43}
	Murray, F. J. \& von Neumann, J.,
	\emph{On rings of operators. IV.}
	Ann. of Math. (2) 44, (1943), pp.~716-–808.
	
\bibitem{niemiec12}
	Niemiec, P.,
	\emph{Elementary Approach to Homogeneous C*-algebras.}
	Rocky Mountain J. Math. 45 (2015), no. 5, pp~1591--1630.	
	
\bibitem{pedersen79}
	Pedersen, G. K.,
	\emph{C*-algebras and their automorphism groups.}
	Academic Press, Inc. [Harcourt Brace Jovanovich, Publishers], London-New York (1979).

\bibitem{rudin91}
	Rudin, W.,
	\emph{Functional Analysis (2nd Ed.).}
	McGraw-Hill (1991).

\bibitem{sakai71}
	Sakai, S.,
	\emph{C*-algebras and W*-algebras.}
	Ergebnisse der Mathematik und ihrer Grenzgebiete, Band 60. Springer-Verlag, New York-Heidelberg (1971).

\bibitem{schroer03}
	Schroer, B.,
	\emph{Pascual Jordan, Glory and Demise and his legacy in contemporary local quantum physics.}
	Unpublished manuscript (2003).
	
\bibitem{simmons83}
	Simmons, G.,
	\emph{Introduction to Topology and Modern Analysis.}
	Robert E. Krieger Publishing Co., Inc., Melbourne, Fl. (1983).

\bibitem{vonneumann40}
	von Neumann, J.,
	\emph{On rings of operators. III.}
	Ann. of Math. (2) 41, (1940), pp.~94–-161.

	
\end{thebibliography}

\end{document}